---
title: "ggsmatr-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ggsmatr-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to create a scatter plot based on the coefficients from (Standardised) Major Axis Estimation fit using the "smatr" package and "ggplot" in R.

Load packages
```{r setup}
library(ggsmatr)
library(smatr)
library(ggplot2)
```

Fit a SMA using smatr package:
```{r}
data(iris)
fit <- sma(Petal.Width ~ Petal.Length+ Species, data=iris)
```

We want to plot the results using ggplot package:

```{r fig1, fig.height = 3, fig.width = 5}
# plot
ggsmatr(data =  iris, groups = "Species", xvar =  "Petal.Length", yvar =
          "Petal.Width", sma.fit =  fit)
```


If we want to change the axis labels, label position and colors we can use the default arguments implemented in ggplot:

```{r fig2, fig.height = 3, fig.width = 5}
#Plot is saved as ggp object, and can be edited using ggplot parameters.
library(ggplot2)
ggp + xlab("x axis name") + ylab("y axis name") +theme(legend.position = "right",
legend.title=element_blank()) + scale_colour_manual(values = c("black", "firebrick", "cornflowerblue"))

```
