---
title: "ggsmatr example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ggsmatr example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4.5
)
```

```{r setup}
library(ggsmatr)
library(ggplot2)
library(smatr)
```

# ggsmatr: Example using the iris dataset

The presence of fitted SMA lines in a ggplot2 scatter plot can be obtained from an object created with `smatr::sma()`. We first read the example data included in the package and we fit a common-slope SMA by group.

```{r fit}
datafile <- system.file("iris.csv", package = "ggsmatr")
df.iris <- read.csv(datafile, encoding = "UTF-8")

fit <- sma(
  Sepal.Length ~ Sepal.Width + Species,
  data = df.iris,
  shift = TRUE,
  elev.test = TRUE,
  alpha = 0.05
)
```

```{r plot-sma}
ggsmatr(
  data = df.iris,
  groups = "Species",
  xvar = "Sepal.Width",
  yvar = "Sepal.Length",
  sma.fit = fit
) +
  theme(legend.position = "top", legend.title = element_blank()) +
  ylab("Sepal.Length") +
  xlab("Sepal.Width")
```

# Confidence ribbons for SMA slopes

Given that SMA lines pass through the group centroid, the presence of a parametric confidence ribbon can be added with `ci = TRUE`. The ribbon is related with the slope confidence intervals already stored in the `sma()` fit (`Slope_lowCI` and `Slope_highCI`), and it is not related with the OLS band of `geom_smooth()`.

The confidence level is the one used when the model is fitted, for example `sma(..., alpha = 0.05)` for 95% intervals. According to Warton et al. (2006, 2012) those limits are obtained by inverting the one-sample test for the SMA slope, and `ggsmatr()` only maps them onto the plot.

Because the fitted line is constrained to pass through the group mean `(xbar, ybar)`, slope and intercept are not independent. The presence of a slope interval is therefore converted to a family of lines through the same centroid:

```r
y_low  <- ybar + Slope_lowCI  * (x - xbar)
y_high <- ybar + Slope_highCI * (x - xbar)
ymin   <- pmin(y_low, y_high)
ymax   <- pmax(y_low, y_high)
```

`pmin()` and `pmax()` are required, because the lower slope produces the higher line when `x < xbar`. The envelope pinches at the group mean and fans out toward the ends of the observed x range. This presents a slope-CI envelope, not a bootstrap prediction band and not a pointwise CI for E[Y | X].

The numerical intervals can be inspected in the `smatr` object:

```{r slope-ci}
fit$groupsummary[, c("group", "Slope", "Slope_lowCI", "Slope_highCI")]
```

```{r plot-sma-ci}
ggsmatr(
  data = df.iris,
  groups = "Species",
  xvar = "Sepal.Width",
  yvar = "Sepal.Length",
  sma.fit = fit,
  ci = TRUE
) +
  theme(legend.position = "top", legend.title = element_blank()) +
  ylab("Sepal.Length") +
  xlab("Sepal.Width")
```

Transparency and smoothness of the ribbon can be controlled with `ci.alpha` and `n`:

```{r plot-sma-ci-opts, eval = FALSE}
ggsmatr(
  data = df.iris,
  groups = "Species",
  xvar = "Sepal.Width",
  yvar = "Sepal.Length",
  sma.fit = fit,
  ci = TRUE,
  ci.alpha = 0.15,
  n = 150
)
```

# References

Warton, D. I., Wright, I. J., Falster, D. S. and Westoby, M. (2006). Bivariate line-fitting methods for allometry. *Biological Reviews* 81, 259–291.

Warton, D. I., Duursma, R. A., Falster, D. S. and Taskinen, S. (2012). smatr 3 – an R package for estimation and inference about allometric lines. *Methods in Ecology and Evolution* 3, 257–259.
